---
title: "RaiseGenic PhenotypicData Vis"
author: "SP"
date: "14 01 2022"
output: 
  html_document: 
    fig_width: 11
    fig_height: 8
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r paths, include=FALSE}
home=getwd()
datafolder=file.path(home, "data", "PhenotypicData")
codebookfolder=file.path(home, "data", "Data_Dict")
genofolder=file.path(home, "data", "VCF_Files")
outputfolder= file.path(home, "output")
```

```{r libraries, include=FALSE}
source(file.path(home, "code", "helper_functions.R"))
load_inst_pkgs("tidyverse", "magrittr", "psych", "labelled")
```

```{r load data}
epi25 <- read_csv(file.path(datafolder, "Epi25_data_merged.csv"), 
                  guess_max = 2000, show_col_types = TRUE)
epi25_codebook <- read_csv(file.path(codebookfolder, "Epi25_merged_labels.csv"))
epi25o <- read_csv(file.path(datafolder, "Epi25_data_merged_outcomes.csv"), 
                   guess_max = 2000, show_col_types = TRUE)
geno_IDs <- read_csv(file.path(genofolder, "samples.txt"),col_names = FALSE) %>% 
  use_series(X1) %>% unique()
```

```{r variable & value labels}
# add variable labels
epi25_labels <- 
  setNames(as.list(epi25_codebook$var_label), epi25_codebook$var_name)
epi25 <- set_variable_labels(epi25, .labels =  epi25_labels, .strict = FALSE)
epi25o <- set_variable_labels(epi25o, .labels =  epi25_labels, .strict = FALSE)
rm(epi25_labels)

# add value labels
# drug labels
var_labels <- list()
var_labels[["drug_labels"]] <- 
  c("Carbamazepine_3", "Lacosamide_11", "Levetiracetam_13", "Valproate_28")
names(var_labels[["drug_labels"]]) <- c("3", "11", "13", "28")
# outcomes
var_labels[["outcome_labels"]] <- 
  c("1_Response", "2_Failure", "3_Unclassified", "4_Unknown")
names(var_labels[["outcome_labels"]]) <- c("1", "2", "3", "4")
# stop reasons
var_labels[["stop_reason_labels"]] <- 
  c("inadequate seizure control", 
    "unacceptable adverse effects", 
    "both inadequate seizure control and unacceptable adverse effects", 
    "remission", 
    "other", 
    "unknown", 
    "NA")
names(var_labels[["stop_reason_labels"]]) <- 
  c("1", "2", "3", "4", "5", "6", "7")
```

TODO: consolidate variables\
- event_name & redcap_event_name

# epi25o outcome data

## enrich data frames
```{r}
# create vars
epi25o %<>% 
  group_by(record_id) %>% 
  mutate(n_events = n(), n_drugs = length(unique(aed_generic_name))) %>% 
  ungroup()

# create subset _ss
epi25o_ss <- epi25o %>% 
  select(record_id, local_identifier, broad_id, aed_generic_name, aed_outcome, 
         stop_reason, sex, yob, start_date, stop_date, still_on_aed, 
         aed_adherent, initiated_as, max_dose, max_dose_serum, appro_aed_trial, 
         adr_on_aed, adr_type, attributed_aed, n_events, n_drugs, 
         starts_with("ethnicity__"), hispanic_status, 
         family_history, fam_first_deg, questionnaire
         # event_name, redcap_event_name
         )

# df for patients with multiple events
epi25o_ss_multiev <- epi25o_ss %>% filter(n_events > 1)
# patients who received the same drug multiple times, but not in different doses
epi25o_ss_samedose <- 
  epi25o_ss %>% 
  group_by(record_id, aed_generic_name) %>% 
  filter(n() > 1, length(unique(max_dose)) == 1)
# exclude those patients for now
excl_pats <- unique(epi25o_ss_samedose$record_id)
```


## variables
```{r}
glimpse(epi25o)
```

TODO - recode 998 etc

## distribution of numeric values
```{r describe}
# epi25 %>% select(where(is.numeric)) %>% describe()
epi25o_ss %>% describe()
```

## available samples
```{r}
epi25o_ss %>% 
  select(record_id, local_identifier, broad_id) %>% 
  summarise(across(.fns = ~length(unique(.))))
n_subj_o <- length(unique(epi25o_ss$record_id))
n_subj_o_g <- length(intersect(unique(epi25o_ss$record_id), geno_IDs))
n_subj_g_no_o <- length(setdiff(geno_IDs, unique(epi25o_ss$record_id)))
```
There are `r n_subj_o` patients with outcome data available.\
`r n_subj_o_g` patients have outcome and Exome data.\
`r n_subj_g_no_o` patients with Exome data have no outcome data.

### drugs per patient
```{r}
epi25o_ss %>% group_by(record_id) %>% slice_head(n = 1) %>% 
  qplot(data = ., x = n_drugs)
```

### number of events & drugs per patient
```{r}
epi25o_ss %>% 
  group_by(record_id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  group_by(n_events, n_drugs) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x = n_events, y = n_drugs)) +
  geom_point(aes(size = count, colour = count)) +
  scale_size(range = c(5, 30)) +
  geom_text(aes(label = count), colour = "white")
```

### inspect patients with multiple events
```{r}
epi25o_ss_multiev %>% 
  filter(n_events == 4) %>% 
  slice_head(n = 4) %>% 
  select(aed_generic_name, aed_outcome, stop_reason, aed_adherent, max_dose, 
         adr_on_aed)
```

### treatment success after drug switch
```{r}
epi25o_ss %>% 
  select(record_id, aed_generic_name, aed_outcome, stop_reason, aed_adherent, 
         max_dose, adr_on_aed, n_drugs, n_events) %>% 
  group_by(record_id) %>% 
  filter(n_drugs > 1) %>% 
  filter(min(aed_outcome) == 1) %>% 
  ungroup() %>% 
  slice_head(n = 4)
```

### patients who received the same drug in different doses
```{r}
epi25o_ss %>% 
  group_by(record_id, aed_generic_name) %>% 
  filter(!c(n() > 1 & NA %in% max_dose)) %>% 
  filter(length(unique(max_dose)) > 1) %>% 
  select(!c(local_identifier, broad_id, sex, yob, still_on_aed, max_dose_serum, 
            adr_type, attributed_aed, starts_with("ethnicity"), hispanic_status, 
            family_history, fam_first_deg)) %>% 
  arrange(record_id, aed_generic_name, start_date)
```

### patients who received the same drug multiple times, but not in different doses
```{r}
epi25o_ss_samedose %>% 
  select(!c(local_identifier, broad_id, sex, yob, still_on_aed, max_dose_serum, 
            adr_type, attributed_aed, starts_with("ethnicity"), hispanic_status, 
            family_history, fam_first_deg))
```
`r length(unique(epi25o_ss_samedose$record_id))` patients received the same drug multiple times with unclear dose relations. Those events are excluded for now.

## ADRs
TODO

# generate outcome dataframe
```{r drug response, warning=FALSE}
# pivot to wide format: 1 row per patient
epi25o_ss_wide <- epi25o_ss %>% 
  filter(!(record_id %in% excl_pats)) %>% 
  pivot_wider(names_from = c(aed_generic_name, max_dose), 
              values_from = c(aed_outcome, stop_reason, start_date, stop_date, 
                              aed_adherent, appro_aed_trial, adr_on_aed, 
                              max_dose_serum, adr_type, initiated_as, 
                              still_on_aed, attributed_aed))

# create 4 main drug outcome columns
epi25o_ss_wide %<>% 
  rowwise(record_id) %>% 
  mutate(aedo_3 = min(c_across(starts_with("aed_outcome_3_")), na.rm = TRUE), 
         aedo_11 = min(c_across(starts_with("aed_outcome_11_")), na.rm = TRUE),
         aedo_13 = min(c_across(starts_with("aed_outcome_13_")), na.rm = TRUE),
         aedo_28 = min(c_across(starts_with("aed_outcome_28_")), na.rm = TRUE)
         ) %>% 
  ungroup() %>% 
  mutate(across(starts_with("aedo_"), ~na_if(., Inf)))

# create smaller drug outcome dataframe
epi25o_ss_dr <- epi25o_ss_wide %>% 
  select(record_id, starts_with("aedo_")) %>% 
  rename("Carbamazepine_3" = aedo_3, "Lacosamide_11" = aedo_11, 
         "Levetiracetam_13" = aedo_13, "Valproate_28" = aedo_28)

write_csv(epi25o_ss_dr, file.path(outputfolder, "aed_responses.csv"))

# check for additional values_from columns (contain multiple values per patient)
# drug_resp %>% group_by(record_id) %>% filter(n() > 1) %>% 
#   summarise(across(.fns = ~length(unique(.)))) %>% ungroup() %>% 
#   summarise(across(.fns = max))
```
TODO: include infos from sanity check cols\
\
- only if appro_aed_trial?
- are there successes (outcome==1) for cases where aed_adherent==0?
- only if stop_reason is 4/remission or 7/NA?
- only if no (adr_on_aed & attributed_on_aed)?


## no patients who didn't receive any drug
```{r all outcomes na}
epi25o_ss_dr %>% 
  filter(is.na(Carbamazepine_3), is.na(Lacosamide_11), is.na(Levetiracetam_13), 
         is.na(Valproate_28))
```

## outcomes for every event
```{r}
table(epi25o_ss$aed_outcome) %>% addmargins()
table(epi25o_ss$aed_outcome) %>% prop.table()*100
```

## outcomes by drug (best outcome per patient)
```{r}
epi25o_ss_dr %>% 
  pivot_longer(cols = !(record_id), 
               names_to = "drug", 
               values_to = "outcome") %>% 
  filter(!is.na(outcome)) %>% 
  mutate(outcome = recode(outcome, !!!var_labels[["outcome_labels"]])) %>% 
  group_by(drug, outcome) %>% 
  summarise(count = n()) %T>% 
  print() %>% 
  ggplot(aes(x = drug, y = outcome)) + 
  geom_point(aes(size = count, colour = count)) + 
  scale_size(range = c(5, 30)) + 
  geom_text(aes(label = count), colour = "white")
```


# epi25 phenotypic data

## ethnicity
```{r}
epi25 %>% 
  summarise(across(starts_with("ethnicity__"), sum, na.rm = TRUE)) %>% 
  t() %>% 
  as_tibble(rownames = "var_name") %>% rename(total = V1) %>% 
  mutate("share_%" = str_c(round(total/sum(total, na.rm = TRUE)*100, 1))) %>% 
  left_join(epi25_codebook)
```

TODO: include hispanic_status in ethnicity\_\_8

## syndrome type ee_syndrome\_, syndrome
```{r}
epi25_codebook %>% filter(str_detect(var_name, "syndrome"))

epi25 %>% 
  summarise(across(starts_with("ee_syndrome__"), sum, na.rm = TRUE)) %>% 
  t() %>% 
  as_tibble(rownames = "var_name") %>% rename(total = V1) %>% 
  mutate("share_%" = str_c(round(total/sum(total, na.rm = TRUE)*100, 1))) %>% 
  left_join(epi25_codebook)

epi25 %>% select(syndrome_focal, syndrome_GGE, syndrome_notes)
```

TODO: recode NAs to 0s for correct counts

```{r}

```

## distribution of seizure types \_seizures

## age of onset by seizure type \_age_onset

## age_last

## comorbidities autism, intellectual_disability + \_degree, psychosis, developmental_delay + \_type,

## family history fam_consanguinity family_history fam_first_deg

## dna available maternal_dna paternal_dna cnv_analysis genetic_testing
